\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{amsmath} 
\newtheorem{lemma}{lemma}
\usepackage{geometry}
\geometry{margin=1in}
\title{Problem Set 1}
\author{Jacob Kohlhepp}
\date{\today}

\begin{document}

\maketitle


The purpose of this homework is to work through our main principal-agent model with uncertainty. The only difference between the model you will solve in this homework and the one in class is that output is not mean 0. You can use the work we did in class as a reference when you get stuck, but please show all steps.

\section*{Setup}
\subsection{Players}
\begin{itemize}
    \item There is a firm (the principal) who is risk neutral (exponential utility with parameter $\theta=0$).
    \item There is a worker (the agent) who is risk averse (exponential utility with parameter $\theta=r\geq 0$).
\end{itemize}
\subsection{Actions}
    \begin{itemize}
    \item Firm chooses a linear wage which depends on effort ($w(e)$) or output ($w(y)$)
    \item After seeing the wage, the worker either accepts or rejects the job.
    \item If they accept, worker chooses effort $e$ at an increasing, convex cost $c(e)=e^2/2$
\end{itemize}
\subsection{Output}
    \begin{itemize}
    \item Output is effort ($e$) plus noise/luck ($\epsilon$): $y=e+\epsilon$ where $\epsilon\sim N(\bar y,\sigma^2)$
    \item Notice that output is now not mean 0.
    \item This implies output is normal with mean $\bar y+e$ and variance $\sigma^2$
\end{itemize}
\subsection{Payoffs}
    \begin{itemize}
    \item If accepted, firm's payoff $\pi$ is expected output minus expected wages: $E[y-w|e]$
    \item If accepted, worker's payoff is expected utility of the wage minus effort cost: $E[u(w) -c(e)|e]$
    \item If rejected, worker has ``outside option" of $\bar u$ and firm has ``outside option" of 0
\end{itemize}

\subsection{Timing}
 The same as in class. The firm proposes a wage schedule, the worker accepts or rejects, the worker exerts effort, output occurs, and then the wage is paid out.


\section{Effort Based Pay}

Suppose the firm can pay based on the workers effort:  $w(e)=\alpha + \beta e$.

\subsection{Questions}
\begin{enumerate}
    \item Is the worker's wage uncertain? Why or why not.
    \item Write the worker's certainty equivalent for a wage with fixed ($\alpha, \beta$) and fixed effort $e$. Hint: use the answer to question 1.
    \item For a fixed wage ($\alpha, \beta$) what level of effort does the agent choose? Hint: your answer should be a function of $\beta$
    \item For what $\alpha, \beta$ does the worker accept the job?
    \item What $\alpha$ will the firm choose, and why? Hint: use the answer to the previous question.
    \item Write the firm's profit. DO NOT simplify or plug anything in.
    \item Simplify the firm's profit. You should get an expression that is a function of only $\beta$ or $e$ (either is fine)
    \item What is the profit-maximizing $\beta$? What is the profit maximizing effort $e$?
    \item What is the profit-maximizing $\alpha$?
\end{enumerate}

\section{Performance-Based Pay}

Suppose the firm can pay based ONLY on output: $w(y)=\alpha + \beta y$ (effort is not observed).
\subsection{Questions}
\begin{enumerate}
    \item Is the worker's wage uncertain? Why or why not.
    \item Write the worker's certainty equivalent for a wage with fixed ($\alpha, \beta$) and fixed effort $e$. Hint: use the answer to question 1.
    \item For a fixed wage ($\alpha, \beta$) what level of effort does the agent choose? Hint: your answer should be a function of $\beta$
    \item For what $\alpha, \beta$ does the worker accept the job? I suggest that you not substitute away $\beta$ yet.
    \item What $\alpha$ will the firm choose, and why? Hint: use the answer to the previous question.
    \item Write the firm's profit. DO not simplify or plug anything in.
    \item Simplify the firm's profit. You should get an expression that is a function of only $\beta$ or $e$ (either is fine)
    \item What is the profit-maximizing $\beta$? What is the profit maximizing effort $e$?
    \item What is the profit-maximizing $\alpha$?
\end{enumerate}

\section{Conclusion}

\subsection{Questions}
\begin{enumerate}
    \item How did allowing for a non-zero average output change the results (hint: we solved the zero mean version in class)?
    \item Provide an intuition or interpretation. Hint: this is deliberately open ended.
\end{enumerate}


\end{document}